url: https://github.com/yaomadrid/marketing-analytics-group-2.git
title: "Marketing Analytics"
output: word_document

1. What is the data problem? What is the final managerial objective?


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

2.Describe the measurement types of each variable




```{r}
summary(car)
1111
```

3. Create a table summarizing the range or variation in each variable. Add statistics (mean,
median, standard deviation, etc.) as you deem necessary.

                                        Actions in loans
```{r}


```


4. How do you handle missing data in this dataset?
```{r}

```


5. Provide histograms/density plots for key variables, such as action taken on loan, gender,
ethnicity, etc.
```{r}
library(dplyr)
ny_hmda_possessed = ny_hmda %>%
group_by(action_taken_name) %>%
summarize(TotalActionTaken = n()) %>%
mutate(PercentageActionTaken = TotalActionTaken/sum(TotalActionTaken) * 100)
library(ggplot2)
ggplot(ny_hmda_possessed, aes(x = action_taken_name,y = PercentageActionTaken))+ geom_bar(stat='identity',colour="white")+labs( title = 'Actions in Loans')

```

6. Create bivariate frequency distributions (tables or plots) for key variables.

7. Discuss what the data patterns indicate, and what this could mean for the probability of
getting a loan approved.
