url: https://github.com/yaomadrid/marketing-analytics-group-2.git
title: "Marketing Analytics"
output: word_document

1. What is the data problem? What is the final managerial objective?


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

2.Describe the measurement types of each variable




```{r}
summary(car)
1111
```

3. Create a table summarizing the range or variation in each variable. Add statistics (mean,
median, standard deviation, etc.) as you deem necessary.

                                       
```{r}


```


4. How do you handle missing data in this dataset?
```{r}

```


5. Provide histograms/density plots for key variables, such as action taken on loan, gender,
ethnicity, etc.

Action in Loans

```{r}
library(dplyr)
ny_hmda_possessed = ny_hmda %>%
group_by(action_taken_name) %>%
summarize(TotalActionTaken = n()) %>%
mutate(PercentageActionTaken = TotalActionTaken/sum(TotalActionTaken) * 100)
library(ggplot2)
ggplot(ny_hmda_possessed, aes(x = action_taken_name,y = PercentageActionTaken))+ geom_bar(stat='identity',colour="white")+labs( title = 'Actions in Loans')+theme(axis.text.x = element_text(angle = 30, hjust = 1))

```



Hustogram of Income
```{r}
library(dplyr)
ny_hmdaIncome <-ny_hmda %>%
  filter(!is.na(hud_median_family_income)) 
plot(density(ny_hmdaIncome$hud_median_family_income), main = "Histogram of income", 
    xlab = "Income", ylab = "Count (#)")
```


Action in Loans by Sex
```{r}
ny_hmdaSex= ny_hmda %>%
  filter(applicant_sex < 3) %>%
  group_by(applicant_sex_name) %>%
  summarize(TotalActionTaken = n()) %>%
  mutate(PercentageActionTaken = TotalActionTaken/sum(TotalActionTaken) * 100)
library(ggplot2)
ggplot(ny_hmdaSex, aes(x = applicant_sex_name ,y = PercentageActionTaken))+ geom_bar(stat='identity',colour="red")+labs( title = 'Actions in Loans')



```

Action in Loans by Race
```{r}
ny_hmdaRace= ny_hmda %>%
  filter(applicant_race_1 < 7 )%>%
  group_by(applicant_race_name_1) %>%
  summarize(TotalActionTaken = n()) %>%
  mutate(PercentageActionTaken = TotalActionTaken/sum(TotalActionTaken) * 100)
ny_hmdaRace 
library(ggplot2)
ggplot(ny_hmdaRace, aes(x = applicant_race_name_1,(angle=30) ,y = PercentageActionTaken))+ geom_bar(stat='identity',colour="red")+labs( title = 'Actions in Loans by race')+theme(axis.text.x = element_text(angle = 30, hjust = 1))

```
1
6. Create bivariate frequency distributions (tables or plots) for key variables. 
Lean_status distribution by actions
```{r}
ny_hmdaLean_status=ny_hmda %>%
group_by(lien_status_name, action_taken_name) %>%
summarize(TotalActionTaken = n()) 
ny_hmdaLean_status
ggplot(ny_hmdaLean_status, aes(x = action_taken_name,y = TotalActionTaken))+ geom_bar(stat='identity',colour="red")+labs( title = 'lean_status distribution by actions')+facet_wrap(~ lien_status_name, scales ="free_y")+theme(axis.text.x = element_text(angle = 30, hjust = 1))
              

```

7. Discuss what the data patterns indicate, and what this could mean for the probability of
getting a loan approved.
